---
title: "MWPC Analysis with eCDF/ePDF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_MWPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## MWPC Analysis: Thresholds, eCDFs, ePDFs, and Customization

Output eCDFs and ePDFs in alignment with FDA Guidance 3. 


```{r output_dir, echo = F}
print.dir = "C:/Users/ciaconangelo/OneDrive - OPEN Health/Documents/misc_R_table_output"
```

## Required R packages
```{r setup}

library(COA34)
# In addition, you'll need the following
library(ggplot2)
library(grid)
library(gridExtra)

```


## Generate data
```{r generate}

  set.seed(12152021)

  sim.out <- COA34::sim_pro_dat_v2(N=1000,
                            number.timepoints = 3,
                            Beta.PRO = NULL,
                            number.of.anchor.groups = 5,
                            polychor.value = 0.7,
                            corr = 'ar1',
                            cor.value = 0.8,
                            #var.values = c(7.136))
                            var.values = c(2))
  dat <- sim.out$dat

 

```



## Implement drop-out

```{r missing}

dat <- COA34::dropout(dat = dat,
               type_dropout  = c('mcar', 'mar', 'mnar'),
               prop.miss = 0.5,
               stochastic.component = 0.2)


```


## Simulate Items
```{r generate items, eval=T}

 # Simulate IRT items
 # use the scores generated as the theta in the IRT model 
  sim.out2 <- COA34::sim_irt_item(dat = dat, J = 5, K = 4, latent.variable = 'Y_comp')
    str(sim.out2)
    dat <- sim.out2$dat

# Score the PRO - just take a simple sum score here:
  dat$PRO.score <- apply(dat[, grep('Item', colnames(dat))], 1, sum)
  
# Create the same PRO score, but with MAR drop-out:
  dat$PRO.score_mar <- dat$PRO.score
  dat$PRO.score_mar[is.na(dat$Y_mar)] <- NA
# Note that you've just set the PRO score to missing wherever the Y_mar variable is missing
  
# Now for the other missing types:
    dat$PRO.score_mcar <- dat$PRO.score_mnar <- dat$PRO.score
    dat$PRO.score_mcar[is.na(dat$Y_mcar)] <- NA
    dat$PRO.score_mnar[is.na(dat$Y_mnar)] <- NA

  aggregate(cbind(PRO.score, PRO.score_mcar, PRO.score_mar, PRO.score_mnar) ~ Time,
                  function(x) mean(x, na.rm = T), 
                  data = dat, 
                  na.action = na.pass)


```



## Compute PGIS delta

```{r pgis_delta}

# You already have the PGIS_bl and PGIS_delta in the generated data,
# you wouldn't have that in a real dataset, so drop that first
dat <- dat[, !(colnames(dat) %in% c('PGIS_bl', 'PGIS_delta'))]
# This makes this more realistic
# use the function below:

dat <- COA34::compute_anchor_delta(dat = dat,
                                   subject.id = 'USUBJID',
                                   time.var = 'Time',
                                   anchor = 'PGIS')

```


## Compute PRO Score delta

```{r score_delta}
dat <- COA34::compute_change_score(dat = dat,
                                   subject.id = 'USUBJID',
                                   time.var = 'Time',
                                   score = c('PRO.score', 'PRO.score_mcar', 
                                             'PRO.score_mar', 'PRO.score_mnar'))

# The function creates variables with the same name plus "_delta":
str(dat)

```

## Compute anchor groups

```{r anchor}
dat <- COA34::compute_anchor_group(dat = dat,
                                   anchor.variable = 'PGIS_delta')


```


## Compute Thresholds

This functions computes a table of all of the meaningful change thresholds. 

```{r thr}

thr <- COA34::compute_thresholds(dat = dat,
                                 anchor.group = 'anchor.groups',
                                 time.var = 'Time',
                                 change.score = 'PRO.score_delta')


```


Output the table to include in the report: 

```{r thr_out, eval = T}
library(R2Word)

R2Word::dump_df_mat_to_file(out = thr,
                            table.title = 'Anchor Group Thresholds',
                            NA.string = '-',
                            decimals = c(2, 2, 0, 1),
                            file.name = 'thr', 
                            print.dir = print.dir)



```


## Compute Proportion Surpassing Threshold

NB, 12.16.21: Not run - not sure how useful this function is.


This function computes the proportion of subjects surpassing a given threshold. So let's say you are looking to estimate meaningful improvement, using a 1-category change on the anchor as the criteria. In this case, you'd use the median PRO score of the anchor group "Improved 1 category" and compute the proportion of subjects with a PRO change score that was that or less (assuming that improvement is negative in PRO change scores).

```{r cap, eval = F}

cap <- COA34::compute_prop_surp(dat = dat,
                                anchor.group = 'anchor.groups',
                                time.var = 'Time',
                                change.score = 'Y_comp_delta',
                                threshold.label = 'Improved_1',
                                mean.or.median = 'median')



library(R2Word)

R2Word::dump_df_mat_to_file(out = cap$anchor.table,
                            table.title = paste0('Proportion Achieving Meaningful Improvement in ', cap$change.score),
                            NA.string = '-',
                            table.footnote = cap$footnote.anchor.table,
                            decimals = c(0, 0, 1),
                            file.name = 'cap', 
                            print.dir = print.dir)


```


## Plot eCDFs and ePDFs
We're going to do this two ways, first the easy way, with a function that automates it. Then we will go and do it by hand, to give you code to customize. 

This way, if the function doesn't give you exactly what you need, you have something to modify and can quickly get what you need.

### Easy Way - Use the function

This is the most expedient approach. You can use this if you have 200 eCDF/ePDFs to compute. 

NB, 12.16.21: function `ggplot2_eCDF` is deprecated, using `ggplot2_eCDF_v2` instead.

```{r plot_funct_1, eval=T}

ecdf <- COA34::ggplot2_eCDF_v2(dat = dat,
                       anchor.group = 'anchor.groups',
                       time.var = 'Time',
                       change.score = 'PRO.score_delta')


epdf <- COA34::ggplot2_ePDF(dat = dat,
                       anchor.group = 'anchor.groups',
                       time.var = 'Time',
                       change.score = 'PRO.score_delta')

# Let's look at them:
plot(ecdf)
plot(epdf)
# Plot both:
gridExtra::grid.arrange(ecdf, epdf, ncol = 1)



```

Note that you can chooose to compute the kernel density for the eCDF if you 
want some smoothed eCDFs. These eCDFs will have lines that don't overlap as much
so it's easier to review and interpret the eCDF. This kernel density is the same
as the one used to plot the ePDFs. 

The default is `smoothed.ecdf = FALSE`. Set it equal to `TRUE`.

```{r plot_funct_2, eval=T}

ecdf.smoothed <- COA34::ggplot2_eCDF_v2(dat = dat,
                       smoothed.ecdf = TRUE,
                       anchor.group = 'anchor.groups',
                       time.var = 'Time',
                       change.score = 'PRO.score_delta')

# Compare the two:
plot(ecdf)
plot(ecdf.smoothed)

```

The plots do not render well in a browser. We should print them out to a file instead.


### Print out the eCDFs

Print out the ggplot2 plots to a .png file. You could do this in a loop if you 
had many figures to print out. 

```{r print_out_1, eval=T, results = 'hide'}

#---------------------------------
# Print out separately:

file.name <- 'Example_eCDF'

png(file = paste0(file.name, '.png'), units="in", width=11, height=8.5, res=300)

  ecdf

dev.off()

#-------------------------------

file.name <- 'Example_eCDF_smoothed'

png(file = paste0(file.name, '.png'), units="in", width=11, height=8.5, res=300)

  ecdf.smoothed

dev.off()

#-------------------------------

  file.name <- 'Example_ePDF'

png(file = paste0(file.name, '.png'), units="in", width=11, height=8.5, res=300)

  epdf

dev.off()
# Note: you can use the paste0 function to use this in a loop
# update anchor_name in the loop
# file.name = paste0('eCDF_ePDF', anchor_name)


#---------------------------------------------------
# Print out together:
  
  file.name <- 'Example_stacked_eCDF_ePDF'

png(file = paste0(file.name, '.png'), units="in", width=11, height=8.5, res=300)

  gridExtra::grid.arrange(ecdf, epdf, ncol = 1)

dev.off()


```


#### Shell Figures

There's an option to return shell figures - this just means the sample sizes
are xx-ed out. This is a nice option if you have to put together a proposal
that contains shell figures and tables. 


```{r plot_shell, eval = T}
ecdf<- COA34::ggplot2_eCDF_v2(dat = dat,
                       shell.table = TRUE,
                       anchor.group = 'anchor.groups',
                       time.var = 'Time',
                       change.score = 'PRO.score_delta')


plot(ecdf)

```

### Custom eCDF

You may want to create a custom plot. There may be one or two scores that are the focus of your report, and for those, you need to make customized modifications.

If you want, you can always pull the R code that is inside these COA34 functions
and customize that code to make it look the way you want. 

However, the nice thing about ggplot2 objects is how you can just add to it. Let's say you want to add a dashed line representing the threshold of deterioration to your eCDF.
That is easy to do: 



```{r plot_custom_ecdf_1, eval = T}

p1 <- COA34::ggplot2_eCDF_v2(dat = dat,
                       anchor.group = 'anchor.groups',
                       time.var = 'Time',
                       change.score = 'PRO.score_delta')


p1 <- p1 +  geom_vline(
  xintercept = thr$Median[thr$`Anchor Group` == 'Deteriorated, 1 category'],
  linetype = 'dashed')

plot(p1)

```


