---
title: "Vignette_Scatterplots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_Scatterplots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Meaningful Change with Small Samples: Scatterplots 

When eCDFs/ePDFs are not feasible to compute, use a scatterplot.
Scatterplots can serve as supplemental information even when eCDFs are computed.


Criteria: When odds ratios in the eCDF approach zero or infinity, it makes more sense to use the scatterplots. In practice, a rule of thumb could be n=5 required in each anchor group for the eCDF to be interpretable. 


TODO: work on developing this criteria further. 

## Required R packages
```{r setup}

library(COA34)
# In addition, you'll need the following
library(ggplot2)
library(grid)
library(gridExtra)

```


## Generate data
```{r generate}

sim.out <- COA34::sim_pro_dat(N=1e3,
                        polychor.value = 0.4,
                        corr = 'ar1',
                        cor.value = 0.8,
                        var.values = c(5))

dat <- sim.out$dat


# You already have the PGIS_bl and PGIS_delta in the generated data,
# you wouldn't have that in a real dataset, so drop that first
dat <- dat[, !(colnames(dat) %in% c('PGIS_bl', 'PGIS_delta'))]
# This makes the resf of this more realistic

```

## Implement drop-out

```{r missing}

dat <- COA34::dropout(dat = dat,
               type_dropout  = c('mcar', 'mar', 'mnar'),
               prop.miss = 0.5,
               stochastic.component = 0.2)


```


## Compute PGIS delta

```{r pgis_delta}
dat <- COA34::compute_anchor_delta(dat = dat,
                                   subject.id = 'USUBJID',
                                   time.var = 'Time',
                                   anchor = 'PGIS')

```

## Compute PRO Score delta

```{r score_delta}
dat <- COA34::compute_change_score(dat = dat,
                                   subject.id = 'USUBJID',
                                   time.var = 'Time',
                                   score = c('Y_comp', 'Y_mcar', 'Y_mar', 'Y_mnar'))


```


## Compute anchor groups

```{r anchor}
dat <- COA34::compute_anchor_group(dat = dat,
                                   anchor.variable = 'PGIS_delta',
                                   number.of.anchor.groups = 5)


```

## Compute Thresholds


```{r thr}

thr <- COA34::compute_thresholds(dat = dat,
                                 anchor.group = 'anchor.groups',
                                 time.var = 'Time',
                                 change.score = 'Y_comp_delta')

```

## Compute Proportion Surpassing Threshold

This function is a little more obscure. 
```{r cap}

cap <- COA34::compute_prop_surp(dat = dat,
                                anchor.group = 'anchor.groups',
                                time.var = 'Time',
                                change.score = 'Y_comp_delta',
                                threshold.label = 'Improved_1',
                                mean.or.median = 'median')


```

# Scatterplots with Anchor-based Meaningful Change

First, adjust the dataframe. The `ggplot2` package is well-suited to long datasets, 
which is the way you should be structuring your data. However, in this case
we have to adjust that. 

```{r dat_prep}
dat.plot <- dat
t4 <- which(dat.plot[ , 'Time'] == 'Time_4')
dat4 <- dat[t4, c('USUBJID', 'Y_comp')]
colnames(dat4) <- c('USUBJID', 'Y_comp_4')
dat.plot <- merge(x = dat.plot, y = dat4, by = 'USUBJID', all = T)

```

Next is to create the x and y limits on the plot. The limits should be
symmetric. Note that real data would likely range from 0 to 100. 

```{r dat_lim}
xlim.max <- round(max(abs(dat.plot$Y_comp)))
use.range <- c(-1*xlim.max, xlim.max)

```

## Prepare Labels for Plot

### Anchor Group Colors
To be used for both dots and lines, and in legends.

```{r dat_1}
unique(dat.plot$anchor.groups)
ag.cols <- c('Deteriorated_2+' = "darkred",
             "Deteriorated_1" = "red",
             'Maintained' = 'blue',
             'Improved_1' = 'green',
             'Improved_2+' = 'darkgreen')
```

### Anchor Group Names - Modify to Include Math Symbols
```{r dat_11}

ag.names <- c('Deteriorated_2+' = '\u2265 2 category deterioration', 
               "Deteriorated_1" = "1 category deterioration",
               'Maintained' = 'Maintained category',
               'Improved_1' = '1 category improvement',
               'Improved_2+' = '\u2265 2 category improvement')#"\u2265" is ≥
```

### Threshold Linetypes 
```{r dat_12}
stat.lty <- c("Mean" = "longdash", "Median" = "dotted")
```


### Anchor Group Labels for Legend
```{r dat_13}
ag.data <- merge(cbind(ag.cols),cbind(ag.names), by = 0)
  colnames(ag.data) <- sub("Row.names", "Anchor Group", colnames(ag.data))
ag.data <- merge(thr, ag.data, by = "Anchor Group")
ag.data$ag.legend <- paste0(ag.data$ag.names, " (N=",ag.data$N,")")
```

This will be merged into a dataframe that contains the threshold values. This
dataframe will then be used in the plot. 

## Dataframe of Thresholds

```{r dat_2}
# Line Plotting data:
line_data <- tidyr::pivot_longer(ag.data,cols = c("Mean", "Median"),names_to = "stat")
line_data$label <- 
  sprintf(paste0(line_data$ag.names,": ",line_data$stat, "=%.2f"),line_data$value)
#TODO: make label a factor to control order in legend
#line_data$label <- factor(line_data$label,
#                           levels = c("≥ 2 category improvement: Mean=-2.28",
#                                      "≥ 2 category improvement: Median=-2.51"))
line_data$linetype <- 
  as.character(factor(line_data$stat,levels=names(stat.lty),labels=stat.lty))
line_data$slope <- 1
line_data$size <- .5#default line size=.5

#add reference line manually
refline <- data.frame("label" = "No Change", "stat" = "", "value" = 0,
                      "ag.cols" = "black", "linetype" = "solid","slope" = 1,"size"=1)
line_data <- merge(line_data, refline, all= T)
```


### Select which thresholds (lines) to plot:

Plot just one threshold line, the median of the "Improved one category" and the median of the "Deteriorated one category" anchor groups:

```{r dat_41}
plot.groups <- c("Improved_1",  "Deteriorated_1")#which anchor groups to include
plot.stats <- c("Median")#which stats to include (Mean and/or Median)
plot_line_data <- line_data[(line_data$'Anchor Group'%in%plot.groups&
                              line_data$stat%in%plot.stats ) |
                              line_data$label=="No Change",]
```

All of the thresholds:
```{r dat_42}
plot_line_data <- line_data # all of the lines
```

## Plot Data using ggplot2 package

```{r dat_43}

# Plot:
p1 <- ggplot(dat.plot, aes(x=Y_comp_bl, y=Y_comp_4, fill = anchor.groups)) +
  theme_minimal() +
  geom_point(shape = 21,size=3, color = "white") + #shape=21 allows a fill, color=white makes no outline
  geom_abline( aes(slope = slope, intercept = value,
                   linetype = label, color=label, size = label),
               data = plot_line_data) +#use data in plot_line_data to get lines
  
  #Line settings: (note 'name' arg is the same for all)
  scale_linetype_manual(
    name = "Threshold",values = setNames(plot_line_data$linetype, plot_line_data$label)) +
  scale_color_manual(
    name = "Threshold",values= setNames(plot_line_data$ag.cols, plot_line_data$label))+
  scale_size_manual(
    name = "Threshold",values= setNames(plot_line_data$size, plot_line_data$label))+
  
  #Point settings:
  scale_fill_manual(
    name = "Anchor Groups", values = ag.cols, 
    labels = setNames(ag.data$ag.legend, ag.data$`Anchor Group`))+
  
  scale_x_continuous( breaks = use.range[1]:use.range[2], limits = use.range) +
  scale_y_continuous( breaks = use.range[1]:use.range[2], limits = use.range) +
  labs( title = 'Baseline & Follow-up Scores by Anchor Group',
        x = 'Score at baseline',
        y = 'Score at follow-up') 


p1



```
