---
title: "Item Response Distribution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_Item_Resp_Dist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Functions to Compute Item Response Distributions

The purpose of this is to show some basic functionality using base R functions. Additional illustrations are available here:  https://cjangelo.github.io/R2Word/articles/introduction.html

These sources are common tables that are rendered as output. 

A much more sophisticated collection of functions and examples is available in the `gtsummary` R package.




```{r output_dir, echo = F}
print.dir = "C:/Users/ciaconangelo/OneDrive - OPEN Health/Documents/misc_R_table_output"
```


## Required R packages
```{r setup}
library(COA34)
library(R2Word)
```


## Generate data
```{r generate}
  set.seed(5122021)
sim.out <- COA34::sim_pro_dat(N=1e3,
                        polychor.value = 0.4,
                        corr = 'ar1',
                        cor.value = 0.8,
                        var.values = c(5))

dat <- sim.out$dat
```

## Implement drop-out

```{r missing}
dat <- COA34::dropout(dat = dat,
               type_dropout  = c('mcar', 'mar', 'mnar'),
               prop.miss = 0.5,
               stochastic.component = 0.2)

```


## Examples

Two useful base R functions (actually in stats package) are `aggregate()` and `xtabs()`.

### Item Response Distributions:
```{r ird1}
out1 <- do.call(data.frame, stats::aggregate(Y_mar ~ Time,
        FUN = function(x) c(sum(!is.na(x)),
                            mean(x, na.rm = T),
                            sd(x, na.rm = T),
                            quantile(x, na.rm = T)),
        data = dat, na.action = na.pass))
# note: be careful of NA handling
colnames(out1) <- c('Time', 'N', 'Mean', 'SD', '0%', '25%', '50%', '75%', '100%')

R2Word::dump_df_mat_to_file(out = out1,
                            decimals = c(0, 2, 3, 2, 2, 2, 2, 2), 
                            table.title = 'Item Response Distribution - Vignette Illustration', 
                            table.footnote = '**All data simulated',
                            file.name = 'ird1', 
                            print.dir = print.dir)
```


### Cross-Tabs:
```{r ct}
out2 <- stats::xtabs( ~  PGIS_delta + Time, data = dat)
out2 <- addmargins(out2, margin = 1)
df <- as.data.frame.matrix(out2)
df <- cbind.data.frame('PGIS_delta' = rownames(out2), df)

R2Word::dump_df_mat_to_file(out = df,
                            decimals = 0,
                            table.title = 'Cross-Table, Vignette Illustration',
                            table.footnote = '**All data simulated',
                            file.name = 'ct',
                            print.dir = print.dir)
```

If you're using a loop, use as.formula() with paste0():
```{r ct2}
x <- 'PGIS_delta'
y <- 'Time'
stats::xtabs( as.formula(paste0('~ ',  x, ' + ', y)), data = dat)
```


### More examples...

Additional examples available here: https://cjangelo.github.io/R2Word/articles/introduction.html


Another great library: `library(dplyr)`. Example:

```{r eg1}
library(dplyr)
dplyr::count(dat, Time, PGIS_bl, PGIS_delta, .drop = F)
```
